---
title: '3.21'
author: ""
date: "November 18, 2017"
output: word_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

"3.21 Generate 10 realizations of length n = 200 each of an ARMA(1,1) process with
phi = .9, theta = .5 and var = 1. Find the MLEs of the three parameters in each case and
compare the estimators to the true values."

```{r import}
packages <- c('stats','pander')
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}

require(stats)
require(pander)

```{r code , echo=TRUE}
#seed(3141593)
phi <- .9
phi.set <- c()
theta <- .5
theta.set <- c()
var <- 1
var.set <- c()

output <- c()

for(i in 1:10){  
  ar2.sim  <- arima.sim(n=200,list(ar=phi,ma=theta,sd=sqrt(var)))
  ar2.fit  <-arima(ar2.sim,order=c(1,0,1)) #p,d,q
  b <- c(i, 'phi,theta,sigma2', paste(round(ar2.fit$coef[1], digits = 2),round(ar2.fit$coef[2],digits=2),round(ar2.fit$sigma2,digits=2),''), paste(phi,theta,var,''))
  phi.set[i] <- round(ar2.fit$coef[1], digits = 2)
  theta.set[i] <- round(ar2.fit$coef[2],digits=2)
  var.set[i] <- round(ar2.fit$sigma2,digits=2)
  output <- rbind(output,b)
}
colnames(output) <- c("Realization", "Parameters","Est. Value", "True Value")
pandoc.table(output)
```

## Visual Comparison


```{r comparison, echo=TRUE}
plot(rep(phi,10), col="red", pch = 15, main="phi")
axis(side=1, at=c(1:10))
points(phi.set, col="blue", pch = 16)
grid(col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

plot(rep(theta,10), col="red", pch = 15, main="theta")
axis(side=1, at=c(1:10))
points(theta.set, col="blue", pch = 16)
grid(col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

plot(rep(var,10), col="red", pch = 15, main="var")
axis(side=1, at=c(1:10))
points(var.set, col="blue", pch = 16)
grid(col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

```


