---
title: '3.42'
author: ""
date: "December 4, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 3.42

3.42 Fit a seasonal ARIMA model of your choice to the unemployment data in
UnempRate. Use the estimated model to forecast the next 12 months.

```{r library, include=TRUE, warning= FALSE}
packages <- c('forecast','astsa','pander')
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), dependencies = TRUE)  
}

require(forecast)
require(astsa)
require(pander)

data(UnempRate)
str(UnempRate)
summary(UnempRate)
```
```
## table 3.1
```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
table.3.1 <- c("","AR(p)","MA(q)","ARMA(p,q)")
table.3.1 <- rbind(table.3.1,c("ACF","Tails off","Cuts off after lag q","Tails off"))
table.3.1 <- rbind(table.3.1,c("PACF","Cuts off after lag p","Tails off q","Tails off"))
pandoc.table(table.3.1)
```

## Data Analytics


```{r analytics, echo=TRUE}
plot.ts(UnempRate)
UnempRate.d.12 <- diff(UnempRate,12)
acf2(UnempRate.d.12)

UnempRate.d.48 <- diff(UnempRate.d.12,48)
acf2(UnempRate.d.48)

UnempRate.fit = arima(UnempRate, order=c(2,1,0), seasonal=list(order=c(0,1,1), period=12))

UnempRate.fit    # to view the results
tsdiag(UnempRate.fit, gof.lag=48)  # diagnostics
```

## Predictions
```{r prediction-1, echo=TRUE}
UnempRate.pr = predict(UnempRate.fit, n.ahead=12)  # forecasts
UnempRate.pr
U = UnempRate.pr$pred + 1.96*UnempRate.pr$se
L = UnempRate.pr$pred - 1.96*UnempRate.pr$se
minx = min(UnempRate,L); maxx = max(UnempRate,U)
ts.plot(UnempRate, col="gray", UnempRate.pr$pred, xlim=c(2005,2018), ylim=c(minx,maxx), main="Unemployment forecast the next 12 months") 
lines(UnempRate.pr$pred, col="red", type="o")
grid(col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)
lines(U, col="blue", lty="dashed")
lines(L, col="blue", lty="dashed")
```

## Predictions
```{r option-2, echo=TRUE}
sarima(UnempRate, 1,0,0,0,1,2,12)
sarima.for(UnempRate, 12, 1,0,0,0,1,2,12)
```