---
title: "Final"
author: ""
date: "December 11, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages <- c('stats','astsa','CombMSC', 'TSA', 'forecast','knitr','pander')
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), dependencies = TRUE)  
}

for(package in packages){
  require(package, character.only = TRUE)
}

set.seed(10)
```

###including necessary components
```{r table3.1, echo=TRUE}
#Table 3.1. Behavior of the ACF and PACF for ARMA models
table.3.1 <- c("","AR(p)","MA(q)","ARMA(p,q)")
table.3.1 <- rbind(table.3.1,c("ACF","Tails off","Cuts off after lag q","Tails off"))
table.3.1 <- rbind(table.3.1,c("PACF","Cuts off after lag p","Tails off q","Tails off"))
pandoc.table(table.3.1)
```
## Question-1
###Let xt be a SARIMA(0, 1, 1)×(1, 0, 1)4:
* i Generate n = 1000 data points using the above model. Plot the time series data.
* ii Use Box-Jenkins method to find an adequate model. Compare with the auto.arima() result.
* iii Write down the algebraic form of your 'best' model and make a 5-ahead forecast (with both plots and predicted numbers).

### i
```{r q1i, include=TRUE}
q1.data.xt <- sarima.Sim(n=1000, period=4, model = list(order= c(0,1,1), ma=.7),seasonal=list(order = c(1,0,1),ar=.6, ma=.8))

q1.data.xt.diff <- diff(q1.data.xt,4)

plot.ts(q1.data.xt.diff,main="ts simulated n=100 period=4",ylim=c(-15,15), col="blue")
grid(col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)
```

### ii
```{r q1ii, include=TRUE}
#seasonality differenced
acf2(q1.data.xt.diff)
#https://www.otexts.org/fpp/8/9

arima.fit <- Arima(q1.data.xt, order=c(4,1,4), seasonal=c(3,0,0), method = "ML")
sarima.fit <- sarima(q1.data.xt,4,1,4,3,0,0,4)

#(p,d,q)x(P,D,Q)4 > (4,1,4)x(3,0,0)4

pred.bj <- sarima.for(q1.data.xt,5, 4,1,4,3,0,0,4)


auto.arima(q1.data.xt, max.p = 5, max.q = 5, max.P = 5, max.Q = 5)
#ARIMA(3,1,3)(1,0,0)[4] 

pred.auto <- sarima.for(q1.data.xt,5, 3,1,3,1,0,0,4)
pred.auto$pred

```

### iii
```{r q1iii, include=TRUE}
plot(c(pred.auto$pred),col="red", pch = 15, type='p' , main="Auto Arima - Box-Jenkins Prediction Comparison", ylab = "sim data", xlab = "time")
text(x=1:5, y=pred.auto$pred + .025, cex = .8, offset = 0, labels=round(pred.auto$pred,2), col="red") 
axis(side=1, at=c(1:5))
points(c(pred.bj$pred), col="blue", pch = 16)
text(x=1:5, y=pred.bj$pred - .025,cex = .8, offset = 0, labels=c(round(pred.bj$pred,2)), col="blue") 
grid(col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)
```

## Question-2

Problem 2. Analyze the data on yearly mean total sunspot number [1700 - now]. The data and information can be found here http://www.sidc.be/silso/datafiles#total.
* i   Use Box-Jenkins method to find an adequate model using the data up to 2010.5 only. Compare with the auto.arima() result.
* ii  Write down the algebraic form of your 'best' model. 
* iii Make a 5-ahead forecast (with both plots and predicted numbers) and compare your forecast with the remainder of the data. Comment.

### data source
* Column 1: Gregorian calendar year (mid-year date)
* Column 2: Yearly mean total sunspot number.
* Column 3: Yearly mean standard deviation of the input sunspot numbers from individual stations.
* Column 4: Number of observations used to compute the yearly mean total sunspot number.
* Column 5: Definitive/provisional marker. A blank indicates that the value is definitive. A '*' 
  symbol indicates that the yearly average still contains provisional daily values and is subject to a possible revision.
  
```{r r2i , echo=TRUE}
sunspot.data <- read.table("D:\\Google Drive\\masters\\MSDA\\STAT5307-Time Series Analysis\\final\\SN_y_tot_V2.0.txt", header = FALSE)
summary(sunspot.data)

sunspot.data.raw = sunspot.data[,1:2]
#317   2
sunspot.data.train <- sunspot.data[1:311,1:2]
sunspot.data.test <- sunspot.data[312:317,1:2]

sunspot.data.train.ts <- ts(sunspot.data.train[,2],start=c(1700),frequency=1)

plot(sunspot.data.train.ts, col="blue")
acf2(sunspot.data.train.ts, 75) 

sunspot.data.train.ts.diff <- diff(sunspot.data.train.ts,12)
acf2(sunspot.data.train.ts.diff)
acf2(diff(sunspot.data.train.ts.diff))
eacf(sunspot.data.train.ts)

#tails off cuts off at Lag 2 AR(2)
```
### Comment
It is one of the properties of ARIMA models, that the AR (autoregressive) and the MA (moving average) terms 
tend to 'eliminate each other out' or 'they work in opposite direction' (if you want to think about this in this way). 
Therefore a model (4,0,4) might be actually very close to (0,0,0) as the 4 AR terms 'eliminate' the effect of 4 MA terms (depends on the coefficients).

Although I am not able to give theoretical foundation for this, it is for practical purposes recommended to fit (if possible) either AR models or MA models, 
because mixing both parts may lead to the above described behavior. It is also a practical rule to not go (if possible) above order of 2 for AR or MA terms.

I have faced similar problems and I also witnessed that if you allow ARIMA models with higher orders (>2) and both (AR and MA) terms, 
it is often the case that models of higher and higher orders will be fitted with better AIC criterion. 
It seems here that the AIC is not fully able to penalize the kind of 'overfitting' that emerges here.

```{r r2iii , echo=TRUE}
arima.fit <- arima(sunspot.data.train.ts,order=c(2,0,0), seasonal=c(0,2,1), method = "ML")
sarima.fit <- sarima(sunspot.data.train.ts,2,0,0,0,2,1,12)

res <- sarima.for(sunspot.data.train.ts,6,2,0,0,0,2,1,12)

res.pred <- c(75.71076, 140.53198, 167.38369, 181.83588, 122.54623,  83.05051)

fit <- auto.arima(sunspot.data.train.ts)
"
ARIMA(2,0,1) with non-zero mean 
Coefficients:
ar1      ar2      ma1     mean
1.4787  -0.7711  -0.1818  79.5908
s.e.  0.0484   0.0438   0.0721   4.0548

sigma^2 estimated as 659.2:  log likelihood=-1449.86
AIC=2909.72   AICc=2909.92   BIC=2928.42
"

autoarima.forecast <- forecast(fit, h = 6)

autoarima.pred <- c(57.41459,  88.96863, 110.55742, 118.15099, 112.73336,  98.86700)

plot(forecast(fit),xlim=c(2000,2016))
text(x=2011:2016, y= autoarima.pred + 2, cex = .6, offset = 0, labels=round(autoarima.pred,2), col="red") 
axis(side=1, at=c(2011:2016))
grid(col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

#plot auto.arima and Box-Jenkins model predictions for next 5 values
plot(x=2011:2016,ylim=c(50,190), y= res.pred,col="red", pch = 15, type='p' , main="Auto Arima - Box-Jenkins Prediction Comparison", ylab = "sim data", xlab = "time")
text(x=2011:2016, y=res.pred + 2, cex = .8, offset = 1, labels=round(res.pred,2), col="red") 
points(x=2011:2016, y= autoarima.pred, col="blue", pch = 17)
text(x=2011:2016, y=autoarima.pred - 2,cex = .8, offset = 0, labels=c(round(autoarima.pred,2)), col="blue") 
grid(col = "lightgray", lty = "dotted", lwd = par("lwd"), equilogs = TRUE)

```

###Comment
Auto arima generated model is not including seasonality component. Box-Jenkins method includes seasonality components
predictions similar
